Stack Safety as a Security Property

### 2020-07-09 main.tex
Clean up keywords
Reorg 3-2
\iflater
Massage 2+3

%% Keywords
%% comma separated list
\keywords{Stack Safety, Micropolicies, Random Testing\bcp{Really??}, RISC-V}  %% \keywords are mandatory in final camera-ready submission

@@

@@

\newcommand*{\add}{\textsc{add}}
\newcommand*{\addi}{\textsc{addi}}
\newcommand*{\sw}{\textsc{sw}}
\newcommand*{\lw}{\textsc{lw}}
\newcommand*{\jal}{\textsc{jal}}
\newcommand*{\jalr}{\textsc{jalr}}
\newcommand*{\rsp}{\textsc{sp}}
\newcommand*{\rra}{\textsc{ra}}
\newcommand*{\rout}{\textsc{out}}

@@

\iflater\bcp{A very small nit: Change

@@

To streamline the example, we use a small instruction set reminiscent of
RISC-V and make a few simplifying assumptions about the organization of
memory.
%
First, since our goal is to formalize stack safety, we ignore the heap and
assume that all variables are stack allocated.
(Integrating a memory safety property for the heap in the style of
\citet{AzevedoDeAmorim17} is a natural extension that we discuss
in~\Cref{sec:future}.)
%
We also assume that the machine is word- (not byte-) addressed, that program
instructions start from location 0, that the machine communicates with the
outside world through a single memory-mapped output port at address 99 (so
a {\tt print} statement will be compiled to a store to this
location), and that the stack begins at location 100 and grows upwards.
% \leo{Explain more?  How to connect to realistic one? Observations? DMA?
%   Interrupt handlers?}\bcp{Yes.}

@@

To make our stack safety definitions precise, we make some simple assumptions about
memory layout; our definitions could easily be changed to be parametric over
these assumptions, at the cost of slightly greater complexity.  Memory 
consists of an instruction region and a stack that grows upwards; for simplicity,
we omit a heap. (Integrating a memory safety property for the heap in the style of
\citet{AzevedoDeAmorim17} is a natural extension that we discuss
in~\Cref{sec:future}.)  The stack has a conventional structure, with
contiguously allocated per-function frames.
The boundary between the frames of a caller and callee is given by the $\rsp$ register
at the time of the call.  The stack can be used to
pass arguments or return results in the usual way, by placing them
at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
definitions say nothing at all about registers (except $\PCname$ and $\rsp$).

Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
policy state to the ordinary machine state\bcp{This may be surprising to
people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
step function that 
refines the machine step function by allowing or disallowing a step to be taken
based on its auxiliary state.   Enforcement works by halting the program before
it can perform an action that would violate the security property; our properties
are naturally therefore \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

This enforcement model is flexible enough to capture any kind of mechanism used
for dynamic enforcement of properties of interest, including the following
representative families of policies.
\apt{I don't really buy this, because we really depend on the ability to define
  variation and rollback. I think we should back off and say: this
  works well for micropolicies, and we conjecture that we can make it work
  for code-altering appraoches (pure software or capability).}

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines like CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including 
the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
(ordinary data) with base and bounds information (policy data),
and their use requires rewriting source code (e.g. to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate. \apt{But we conjecture it can be done...?}

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the compiler to
(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
canaries~\citep{Cowan+98}. \apt{not sure what to say here}
%These approaches are represented as policies that
%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
%being produced by the compiler.) The set of policy states is again the unit type
%and the step function constant. Property enforcement comes from the fact that
%only ill-formed states can induce a trace that violates the property.

\rb{TODO In a way, in these schemes the policy state is built in as part of the
  program state and transparent to the hardware, although it could be simulated
  by an overlay where enforcement is represented at the policy level.}

>>>>

%% Keywords
%% comma separated list
\keywords{Stack Safety, Micropolicies}  %% \keywords are mandatory in final camera-ready submission

@@

\newcommand*{\rsp}{\textsc{sp}}

To make our stack safety definitions precise, we make some simple assumptions about
memory layout; our definitions could easily be changed to be parametric over
these assumptions, at the cost of slightly greater complexity.  Memory 
consists of an instruction region and a stack that grows upwards; for simplicity,
we omit a heap. (Integrating a memory safety property for the heap in the style of
\citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
in~\Cref{sec:related}.)  The stack has a conventional structure, with
contiguously allocated per-function frames.
The boundary between the frames of a caller and callee is given by the $\rsp$ register
at the time of the call.  The stack can be used to
pass arguments or return results in the usual way, by placing them
at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
definitions say nothing at all about registers (except $\PCname$ and $\rsp$).

Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
policy state to the ordinary machine state
%\bcp{This may be surprising to people who are not thinking about micropolicies.\apt{added para below}}
and a partial policy step function that refines the machine step
function by allowing or disallowing a step to be taken based on its
auxiliary state.  Enforcement works by halting the program before it
can perform an action that would violate the security property; our
properties are naturally therefore \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

This enforcement model is flexible enough to capture
micropolicies~\citep{pump_oakland2015}, and we believe that it can
also capture code-altering approaches (such as software-based
or capability-based techniques).

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines like CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including 
the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
(ordinary data) with base and bounds information (policy data),
and their use requires rewriting source code (e.g. to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate. \apt{But we conjecture it can be done...?}\leo{Just say that?}

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the
compiler to (e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or
to insert stack canaries~\citep{Cowan+98}. Such approaches should
still satisfy our formal notion of stack safety with a trivial null
policy, but only for programs produced by the compiler.
\apt{not sure what to say here}
\leo{Something like this?}
%These approaches are represented as policies that
%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
%being produced by the compiler.) The set of policy states is again the unit type
%and the step function constant. Property enforcement comes from the fact that
%only ill-formed states can induce a trace that violates the property.

@@

\newcommand*{\add}{\textsc{add}}
\newcommand*{\addi}{\textsc{addi}}
\newcommand*{\sw}{\textsc{sw}}
\newcommand*{\lw}{\textsc{lw}}
\newcommand*{\jal}{\textsc{jal}}
\newcommand*{\jalr}{\textsc{jalr}}
%\newcommand*{\rsp}{\textsc{sp}}
\newcommand*{\rra}{\textsc{ra}}
\newcommand*{\rout}{\textsc{out}}

@@

\ifaftersubmission\bcp{A very small nit: Change

@@

To streamline the example, we use a small instruction set reminiscent of
RISC-V and make a few simplifying assumptions about the organization of
memory.
%
First, since our goal is to formalize stack safety, we ignore the heap
and assume that all variables are stack allocated. As noted in the
previous section, the extension to account for memory safety is
discussed in~\Cref{sec:related}.
%
We also assume that the machine is word- (not byte-) addressed, that program
instructions start from location 0, that the machine communicates with the
outside world through a single memory-mapped output port at address 99 (so
a {\tt print} statement will be compiled to a store to this
location), and that the stack begins at location 100 and grows upwards.
% \leo{Explain more?  How to connect to realistic one? Observations? DMA?
%   Interrupt handlers?}\bcp{Yes.}

@@
